---
title: "2022 Week 4 Lab"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

Import Data

```{r}
library(igraph)
### The nodes file contains the ID codes of doctors and then considerable information about them - their specialty, number of patients of certain types, network data etc

###  The edges file is a list of teh connections, with an edge weight that equals the number of shared patients

data_dir <- "Data"
LAedges<-read.csv(file.path(data_dir, "LAedges.csv"))
LAnodes<-read.csv(file.path(data_dir, "LAnodes.csv"))

### Remove the first column of the edgelist
LAedges <- LAedges[ -c(1) ]

```

Create the Network and Plot 
```{r}
LA.doctors <- graph_from_data_frame(LAedges, directed=FALSE, vertices = LAnodes)

#preppat is number of prescriptions so we're sizing by that (Scaled)
plot(LA.doctors,
     vertex.size=V(LA.doctors)$preppat/10,
     vertex.label=NA)

```


Filter to Drs in SM
```{r}
SM.doctors<-induced.subgraph(LA.doctors, which(V(LA.doctors)$ZIP %in% c("90401","90402", "90403", "90404", "90405"  )))

plot(SM.doctors,
     vertex.size=V(SM.doctors)$preppat/10,
     vertex.label=NA)

```

Filter to the largest component

```{r}
### extract only the connected component, i.e. ignore all of the doctors around the edge that aren't linked to other doctors in this dataset

# extract the components
SM.doctors.components <- clusters(SM.doctors)
# which is the largest component
SM.doctors.max <- which.max(SM.doctors.components$csize)
# get the subgraph correspondent to just the giant component
SM.doctors.g <- induced.subgraph(SM.doctors, which(SM.doctors.components$membership == SM.doctors.max))


plot(SM.doctors.g,
     vertex.size=V(SM.doctors.g)$preppat/5,
     vertex.label=NA)

```


Community Detection

```{r}

### I'll go over this in the upcoming lecture, but igraph has several ways to group networks into communities,in which each node is assigned to one community based on a particular algorithmic strategy

## example of fastgreedy community detection
SM.doctors.comms <- fastgreedy.community(SM.doctors.g)
V(SM.doctors.g)$community <- membership(SM.doctors.comms)


SM.doctors.g.comms2 <-spinglass.community (SM.doctors.g)   ## the spinglass algorithm can only work on connected graphs
V(SM.doctors.g)$community2 <- membership(SM.doctors.g.comms2)


# number of communities
length(SM.doctors.comms)
length(SM.doctors.g.comms2)

# size of communities
sizes(SM.doctors.comms)
sizes(SM.doctors.g.comms2)


# crossing edges: Which edges cross multiple communities?
crossing.edges<-crossing(SM.doctors.comms, SM.doctors.g)  ### not terribly rewarding in a large network

table(crossing.edges)

crossing.edges2<-crossing(SM.doctors.g.comms2, SM.doctors.g)

table(crossing.edges2)

# plot communities with shaded regions
plot(SM.doctors.g.comms2, # communities
     SM.doctors.g, # graph
     vertex.size=V(SM.doctors.g)$preppat/10,
     vertex.label=NA)


```


Filter to one community

```{r}

### you can subset the graph to look at particular communities as well. Use sizes(SM.doctors.comms.g.2) to figure out the community with the biggest size. It's actually 14 with 22 nodes

SM.doctors.g.2.community1 <- induced.subgraph(SM.doctors.g, which(V(SM.doctors.g)$community2 == 14))

# induced.subgraph(SM.doctors, which(SM.doctors.components$membership == SM.doctors.max))

SM.doctors.g.2.community1

plot(SM.doctors.g.2.community1,
     vertex.size=V(SM.doctors.g)$preppat*10,
     vertex.label=NA)

```



# Assignment Notes

Lab: Physician data  

Baseline Assignment:  

1) Load and orient yourself to a dataset of physician patient sharing

2) Compile a table of summary statistics about the network structure (of a few communities)

3) Identify key doctors in the network (or community)

<!-- #### explore the data -->
<!-- ### try things like making the nodes different colors depending on whether they are PrEP prescribers,  -->
<!-- ### perhaps size them by the e number of PrEP, HIV or Herpes patients they have -->
<!-- ### perhaps see if there are interesting patterns that emerge if you examine different physician specialty codes? -->

