---
title: "Lab 3"
date: "`r Sys.Date()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(igraph)
library(tidyverse)
```

## Load Data

```{r}
data_dir <- 'Data'

directors <- read_csv(file.path(data_dir, 'Dir2020raw.csv'))
firms <- read_csv(file.path(data_dir, 'Firm_LU.csv')) %>% distinct(BvD, .keep_all=TRUE)
people <- read_csv(file.path(data_dir, 'People_LU.csv')) %>% distinct(UCI, .keep_all=TRUE)

glimpse(directors)
glimpse(firms)
glimpse(people)
```

## Clean Up the Dataset

`directors` is our "master" edge list...each row describes a _relationship_ between a firm and a person employed in some sort of directorship role. We can say that the people _lead_ firms to characterize the bipartite relationship. `firms` contains characteristics of the firms; one row represents one firm. Similarly, `people` contains characteristics of the individual; one row represents one person. This data looks like it came from a relational database, where `BvD` represents the key that relates firms to directorships and `UCI` is the key that relates people to directorships.

```{r}
small_dirs <- sample_n(directors, 5000, replace=F)
```

```{r}
dirs_el <- small_dirs %>% select(UCI, BvD, DMtitle)
dirs_net <- graph_from_data_frame(dirs_el)
V(dirs_net)$type <- V(dirs_net)$name %in% people$UCI # True is person, False is firm

vertex_attr_names(dirs_net)

bipartite_proj_list <- bipartite_projection(dirs_net)
bipartite_proj_list

bipart_firms <- bipartite_proj_list$proj1
bipart_people <- bipartite_proj_list$proj2

bipart_people_df <- igraph::as_data_frame(bipart_people, 'both')
bipart_people <- graph_from_data_frame(bipart_people_df$edges, vertices = bipart_people_df$vertices %>% left_join(people, by=c("name" = "UCI")), directed=F)

V(bipart_people)$color <- as.factor(V(bipart_people)$ISO)

plot(bipart_people, vertex.label=NA, vertex.size=sqrt(degree(bipart_people))+1, layout=layout_with_dh)
```


## Lab Code

### Set source and target for the graph

```{r}
# Rename employee ID 
colnames(RAND_work)[1]<-"EMPID"

###  Make source and target columns (not really required for igraph)
RAND_work$Source <- RAND_work$EMPID
RAND_work$Target <- RAND_work$PTN

head(RAND_work)

```

### Reorder Columns 
```{r}
####  Reorder columns so that the source and target ids are in the first two columns
RAND_work <-RAND_work[,c(8,9,1,4,2,3,5,6,7)]

head(RAND_work)
```


### Filter the Dataset 

```{r}
#  Filter the dataset so that people working for more than 10 days are included and only charges from 2018 are used
RAND_work_c <-subset(RAND_work, TOTAL_HOURS > 80)
RAND_work_c.18 <-subset(RAND_work_c, FY > 2017)
head (RAND_work_c.18)
```

## Create a Nodelist

```{r}
#### Create a node attribute table for people and select only people attributes
People_nodes <-RAND_work_c.18
People_nodes.d <- People_nodes[c(3,7,8)]

### remove duplicates 
People_nodes.d <- dplyr::distinct(People_nodes.d, EMPID, .keep_all = TRUE)


```

## Create a graph 
```{r}
#### remove everything but source and target
RAND.18 <- RAND_work_c.18[c(-3,-4, -5, -6, -6, -7, -8, -9)]

#### make a bimartite network
RAND.net.18 <- graph_from_data_frame(RAND.18)
plot(RAND.net.18)

RAND.net.18
```
## Check if it will project properly as a bipartite network
```{r}
### create a "type" column to identify the two types of nodes
V(RAND.net.18)$type <-V(RAND.net.18)$name %in% RAND_work_c.18$Source # True is person, False is ptn
is_bipartite(RAND.net.18) # checks whether the graph has the graph has a vertex attribute called type which is necessary for the projection
```


```{r}
### take the two projections of the network
RAND.list<-bipartite_projection(RAND.net.18, type=V(RAND.net.18)$type, multiplicity = TRUE) # Multiplicity to double count individuals connected through more than one ptn 
RAND.bipar.1<-RAND.list$proj1  ## PTNs
RAND.bipar.2<-RAND.list$proj2  ### People

RAND.list
```

```{r}
###  Select the People network and add vertex attributes
#### Department
People2018.network<-set_vertex_attr(RAND.bipar.2,
                                    name="dept",
                                    index = V(RAND.bipar.2),
                                    value=People_nodes.d$DEPT_10_10_FY)
#### Title
People2018.network<-set_vertex_attr(People2018.network,
                                    name="title",
                                    index = V(RAND.bipar.2),
                                    value=People_nodes.d$TITLE_10_10_FY)

V(People2018.network)[V(People2018.network)$dept == "BPS"]$color <- "green"
V(People2018.network)[V(People2018.network)$dept == "DPS"]$color <- "blue"
V(People2018.network)[V(People2018.network)$dept == "ESS"]$color <- "orange"
V(People2018.network)[V(People2018.network)$dept == "EAS"]$color <- "red"


#### Behold!
plot(People2018.network,
     vertex.color=V(People2018.network)$color,
     vertex.size=4,
     vertex.label=NA)

```

Let's look at the network of senior behavioral and social scientists 
```{r}


People2018.network.1<-induced.subgraph(People2018.network, which(V(People2018.network)$title=="Behavioral/Soc Sci, Sr"))

plot(People2018.network.1,
     vertex.color=V(People2018.network.1)$color,
     vertex.size=8,
     vertex.label=NA)

```

Let's compare it to the network of senior political scientists

```{r}


People2018.network.2<-induced.subgraph(People2018.network, which(V(People2018.network)$title=="Political Sci, Sr"))

plot(People2018.network.2,
     vertex.color=V(People2018.network.1)$color,
     vertex.size=8,
     vertex.label=NA)

```

Let's look at which type of worker has the highest average degree centrality

```{r}
V(People2018.network)$deg<-degree(People2018.network, v = V(People2018.network), mode = c("all"),  normalized = FALSE)

data.frame(deg=V(People2018.network)$deg,
           title=V(People2018.network)$title) %>% 
  group_by(title) %>% 
  summarise(mean.deg=mean(deg)) %>% 
  arrange(desc(mean.deg)) 
  
```