---
title: "Lab 3"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(igraph)
library(tidyverse)
```

## Load Data

```{r}
data_dir <- 'Data'

directors <- read_csv(file.path(data_dir, 'Dir2020raw.csv'))
firms <- read_csv(file.path(data_dir, 'Firm_LU.csv')) %>% distinct(BvD, .keep_all=TRUE)
people <- read_csv(file.path(data_dir, 'People_LU.csv')) %>% distinct(UCI, .keep_all=TRUE)

glimpse(directors)
glimpse(firms)
glimpse(people)
```

## Clean Up the Dataset

`directors` is our "master" edge list...each row describes a _relationship_ between a firm and a person employed in some sort of directorship role. We can say that the people _lead_ firms to characterize the bipartite relationship. `firms` contains characteristics of the firms; one row represents one firm. Similarly, `people` contains characteristics of the individual; one row represents one person. This data looks like it came from a relational database, where `BvD` represents the key that relates firms to directorships and `UCI` is the key that relates people to directorships.

```{r}
small_dirs <- sample_n(directors, 20000, replace=F)
```

```{r}
dirs_el <- small_dirs %>% select(UCI, BvD, DMtitle)
dirs_net <- graph_from_data_frame(dirs_el)
V(dirs_net)$type <- V(dirs_net)$name %in% people$UCI # True is person, False is firm

vertex_attr_names(dirs_net)

bipartite_proj_list <- bipartite_projection(dirs_net)
bipartite_proj_list

bipart_firms <- bipartite_proj_list$proj1
bipart_people <- bipartite_proj_list$proj2

bipart_people_df <- igraph::as_data_frame(bipart_people, 'both')
bipart_people <- graph_from_data_frame(bipart_people_df$edges, vertices = bipart_people_df$vertices %>% left_join(people, by=c("name" = "UCI")), directed=F)

V(bipart_people)$color <- as.factor(V(bipart_people)$ISO)

plot(bipart_people, vertex.label=NA, vertex.size=sqrt(degree(bipart_people))+1, layout=layout_with_dh)
```