---
title: "Lab 1"
author: "Ignacio Lara"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Building and Plotting the Network

1. Install and load the packages readr and igraph

```{r message=FALSE, warning=FALSE}
### downloads and installs new features from the internet.  You only need to run these commands once.

# install.packages("readr")
# install.packages("igraph")

## You need to load the particular fatures (packages) that you'll use for a given session
library(readr)
library(igraph)
library(tidyverse)
```

2. Import nodes and edges as separate data frames using *read_csv* from readr

```{r}
## here, i'm importing the cvs files that contain the raw data entered through excel
## the read_csv function is part of the readr package
data_dir <- "Data"
airports <- read_csv(file.path(data_dir, "airportnodes.csv"))
routes <- read_csv(file.path(data_dir, "routesel.csv"))
```
### Explore the Data
#### Airports data
First, let's just get a feel for the columns, their types, and some of their values using `glimpse()`. Then, we'll look at missing values.
```{r}
glimpse(airports)
colSums(is.na(airports))
```
It looks like we have some duplicated information, probably from a merge between different datasets. For example, `id`, `ID`, and `IATA` all seem to refer to the International Air Transport Association (IATA) 3-letter airport codes. Let's check the number of times those columns match:

```{r}
summary <- airports %>% group_by(id) %>% summarize(across(c('ID','IATA'), n_distinct))
summary
summary %>% summarize(across(-id, .fns = max))
```
The first summary table shows the number of unique values of `ID` and `IATA` for each value of `id`. The second summary table finds the maximum number of unique values. Since the maximum number is 1, we can be confident that the `id`, `ID`, and `IATA` columns always match. 

Since `id` and `ID` have 0 missing values, but `IATA` has 750 `NA`s, there are some airports contained in the dataset(s) that used `id` / `ID` which are missing from the dataset that used `IATA`. Let's keep the `id` column to preserve these airports, and explore what the rows with missing `IATA` are like.
```{r}
airports %>% filter(is.na(IATA))
nodes <- airports %>% select(-c('ID','IATA'))
glimpse(nodes)
```
Based on the `filter` we used, it seems that the `airportID:source` columns (all columns from `airportID` to `source`) came from a different data set, and the airports with missing values from `IATA` are just absent from this second dataset (because all other columns in `airportID:source` for these rows are also `NA`). 

To be consistent with which `latitude` and `longitude` we use across the airports in our dataset, let's stick with the ones that came from the first dataset, and drop the ones from the second dataset (`longitude.y` and `latitude.y`).

Finally, since we are only really interested in the graph of flight routes, where nodes are departure/destination airports and arcs are (directed) routes operated by airlines between airports, we don't need a lot of the extra detail like the ICAO code, time zone, altitude, etc.
```{r}
nodes <- nodes %>% select(id:country) %>% select(-X1)
glimpse(nodes)
```
#### Routes data

Now we'll follow a similar process with the `routes` dataset.
```{r}
glimpse(routes)
colSums(is.na(routes))
```
There are also some ambiguous names here, possibly indicative of repeated data. What does `country` refer to?
```{r}
routes %>% group_by(Source) %>% summarize(countries = n_distinct(country))
routes %>% group_by(Target) %>% summarize(countries = n_distinct(country))
routes %>% group_by(airline) %>% summarize(countries = n_distinct(country))
```
So `country` cannot refer to the country of the departure or arrival airport, since the same airport code has multiple unique values of `country`. It seems that `country` is then related to `airline`, since each `airline` has only a handful of values for `country`, and most `airline`s have only 1 `country`:
```{r}
mult_airlines <- routes %>% group_by(airline) %>% summarize(countries = n_distinct(country)) %>% filter(countries > 1) %>% .$airline
mult_airlines

routes %>% select(airline, airline_name, country) %>% filter(airline %in% mult_airlines) %>% unique()
```
It looks like some two-letter airline codes are repeated, perhaps because airlines registered in different countries are allowed to have the same code, or maybe some older airlines with routes in the data have since gone out of business or changed codes with another airline. Again, since we only really care about which airport pairs are served by _any_ airline. In total, 113 of the 568 airline codes are associated with more than 1 country. Overall, though, I'm not concerned about this data issue.
```{r}
arcs <- routes %>% select(Source, Target)
glimpse(arcs)
arcs <- arcs %>% count(Source, Target, name='weight')
glimpse(arcs)
```
The last step counts the number of times the unique airport pair occurs in the `routes` data. Since the data is directed, each one-way flight between two airports would be listed, i.e. $AAA \rightarrow BBB$ and $BBB \rightarrow AAA$ would be listed separately.

3. Combine the nodelist and edgelist into a directed graph using *graph_from_data_frame* 

```{r}
g <- graph_from_data_frame(arcs)
```

4. Plot the graph using *plot*. 

```{r}
plot(g, edge.arrow.size=0.2, vertex.size=1, vertex.label.cex=0.4, edge.width = E(g)$weight, vertex.label.family='Helvetica', layout=layout_with_lgl)

```
```{r}
V(g)$latitude <- airports$latitude[match(V(g)$name, airports$id)]
V(g)$longitude <- airports$longitude[match(V(g)$name, airports$id)]

latlong = matrix(c(V(g)$longitude, V(g)$latitude), ncol=2)
```


```{r}
plot(g, edge.arrow.size=0.05, vertex.size=1, vertex.label=NA, edge.width = 0.1*E(g)$weight, vertex.label.family='Arial', layout=latlong, rescale=FALSE, asp=0, xlim=range(V(g)$longitude), ylim=range(V(g)$latitude))
```


```{r}
library(tidyverse)
world <- map_data("world")

coords <- nodes %>% select(id:longitude)
el <- as_edgelist(g) %>% as_tibble
edges <- el %>% full_join(coords, by=c('V1' = 'id')) %>% 
                full_join(coords, by=c('V2'='id'), suffix=c('.from', '.to')) %>% filter(V1 != V2)

ggplot() +
  geom_map(data=world, map=world, aes(long, lat, map_id=region), color='gray', fill='white', size=0.05) +
  geom_point(data=nodes, aes(x=longitude, y=latitude), size=0.1, color='lightblue') +
  geom_curve(data=edges, aes(x=longitude.from, xend=longitude.to,
                               y=latitude.from, yend=latitude.to),
               size=0.5, alpha=0.02, color='darkgreen')#+
  #coord_map("ortho", orientation=c(30, 10, 0))
```